Total_SC

### 2020-02-14 doc/Results.rmd
starting total SC analysis

@@

                       'PWD'= 'red2',
                       'SKIVE'= 'coral1','KAZ'= 'coral4', 'TOM'= 'indianred1','AST' = 'red4',
                       'CZECH'= 'indianred4',

@@

@@

@@

@@

```{r show.SC.length, echo=FALSE, warning=FALSE}
#for now remove the 
#Curated_BivData.99 <- Curated_BivData[Curated_BivData$hand.foci.count #<5,]
#CHECK THAT THE ABOVE TOO MANY FOCI BIVALENTS DON"T EXSIS

#add boxplots for mouse level means
#add second x axis, hand.foci.count

#other_thing22 <- ggplot(data = Curated_BivData.99, aes(y = chromosomeLength, x = hand.foci.count, fill=as.factor(mouse)),color=strain ) + geom_jitter() + facet_wrap(sex~subsp, scales = "free")+theme(legend.position="none")+scale_color_manual(values=colors_of_strains)

#SC.lengths.1CO <- ggplot(Curated_BivData[Curated_BivData$hand.foci.count == 1,], aes(y = chromosomeLength, x=strain, fill=sex, color=strain))+geom_boxplot(aes(alpha=0.5))+geom_jitter(aes(alpha=0.2))+  ggtitle("1COs male and females") +ylim(c(30,250))+scale_color_manual(values=colors_of_strains)
Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]

#nested boxplots
SC.length_Nested_boxplots <- ggplot(Curated_BivData99, aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(alpha=0.5))+ggtitle("SC Lengths by Chrm Class")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))

@@

@@

@@

write.table(round(summry.glm.Q1_IFD.ABS.M3$coefficients,6), "~./MLH1repo/summry.glm.Q1_IFD.ABS.M3.csv", sep="\t",row.names = TRUE)

@@

write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), "~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)

@@

long.biv data annd short bivdata

Focus on comparing the differences between long.biv (and short biv once I make that dataset)

@@

>>>>

load(file="~./MLH1repo/data/total_SC_30.12.19.RData")

#skeletonize
SC.skel = read.csv("~./SC_skeletonize/data/SCskel_output_nov19.csv", header = TRUE, strip.white = TRUE)

@@

                       'PWD'= 'red2',
                       'SKIVE'= 'coral1','KAZ'= 'coral4',

                       'TOM'= 'indianred1','AST' = 'tan2',
                       'CZECH'= 'brown1',

@@

color_chrm_class <- c( '0' = "#00BFC4", '1' = "#7CAE00", '2' = "#F8766D", '3' = "#C77CFF")

@@

```{r totSC_setup, echo=FALSE}
#SC.skel
#"anon.batch" "rand_name"  "bin_size"   "skel_size"

#MERGEING MLH1 with SKEL
MLH1.merge.Skel <- merge(original_DF, SC.skel, by.x = "Random.Name", by.y = "rand_name", all= F)
OG.MLH1.merge.Skel <- MLH1.merge.Skel #raw version of DF

MLH1.merge.Skel <- add_mouse(MLH1.merge.Skel)
MLH1.merge.Skel <- add_strain(MLH1.merge.Skel)
MLH1.merge.Skel <- add_sex(MLH1.merge.Skel)
MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)
MLH1.merge.Skel <- add_category(MLH1.merge.Skel)
MLH1.merge.Skel <- add_subsp(MLH1.merge.Skel)

MLH1.merge.Skel <- MLH1.merge.Skel %>% filter( X == "") %>% filter(REDO.crop == "no")

MLH1.merge.Skel$bin_size <- as.numeric(MLH1.merge.Skel$bin_size)
MLH1.merge.Skel$skel_size <- as.numeric(MLH1.merge.Skel$skel_size)

#table(MLH1.merge.Skel$category)

!Note! the most recent tot.sc csv was missing ~2 batches... rerun

```{r tot.sc.viz4outliers, echo=FALSE}

tot.sc <- ggplot(data= MLH1.merge.Skel, aes(y = bin_size, x=mouse))+ geom_point()
tot.sc2 <- ggplot(data= MLH1.merge.Skel, aes(y = skel_size, x=mouse))+ geom_point()

#seems like the outliers have been removed
```

```{r tot.skel.table, echo=FALSE}

totSC.mouse.av.table <- ddply(.data=MLH1.merge.Skel, 
                 .(mouse),
                 summarize, 
                 ncells = length(unique(Original.Name)),
                 nmice = length(unique(mouse)),
                # mean.MLH1 = mean(adj_nMLH1.foci, na.rm = TRUE),
                # var.MLH1 = var(adj_nMLH1.foci, na.rm = TRUE),

                 mean.skel = mean(skel_size, na.rm = TRUE),
                #se for skel

                 mean.bin =  mean(bin_size, na.rm = TRUE)

totSC.mouse.av.table <- add_strain(totSC.mouse.av.table)
totSC.mouse.av.table <- add_sex(totSC.mouse.av.table)
totSC.mouse.av.table <- add_category(totSC.mouse.av.table)
totSC.mouse.av.table <- add_subsp(totSC.mouse.av.table)

#unorder the factors
totSC.mouse.av.table$subsp <- factor(totSC.mouse.av.table$subsp, ordered = FALSE)
totSC.mouse.av.table$sex <- factor(totSC.mouse.av.table$sex, ordered = FALSE)
totSC.mouse.av.table$strain <- factor(totSC.mouse.av.table$strain, ordered = FALSE)

#add rec group
totSC.mouse.av.table$Rec.group <- ifelse(grepl("PWD male", totSC.mouse.av.table$category), 1, 
                              ifelse(grepl("MSM male", totSC.mouse.av.table$category), 1,
                                ifelse(grepl("SKIVE male", totSC.mouse.av.table$category), 1, 0)))

OG.totSC.mouse.av.table <- totSC.mouse.av.table

totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")

tot.sc99 <- ggplot(data= totSC.mouse.av.table, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(sex~subsp, scales="free")

tot.sc77 <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ylim(c(0, 800))+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

tot.sc_bin <- ggplot(data= OG.totSC.mouse.av.table[OG.totSC.mouse.av.table$sex == "male",], aes(y = mean.bin, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())  +ylim(c(0, 3000))

#there should be tom

```
From this figure of the male mouse average of total sc measures, Confirm the general pattern of more SC area for the higher rec strains, compared to the low. But there are some exceptions.

the low rec musc strains which are higher are czech (1) and ast (2) -- the PWD mouse which is lower 24sep14_PWD_m1

INSERT TABLE FOR THE TOTAL SC means ect

The difference between

bin is short for binary, it reports the binarized area of the SC

skel -- takes the skeletonized midline of the bin objects

@@

In order to test one of the major/known correlates with genome wide recombiation rate variation, we compared the average
```{r Q1.show.SC.length, echo=FALSE, warning=FALSE}

Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]

#adjust order of plot here
Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
  "WSB","G","LEW","PERC",  
"PWD", "SKIVE",  "KAZ",
     "MSM", "MOLF",   
                     "CAST",
                         "SPRET","SPIC" ), order=T )

#nested boxplots, pooled Bivalents

Q1.SC.length_Nested_boxplots_Dom <- ggplot(Curated_BivData99[Curated_BivData99$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none") 
Q1.SC.length_Nested_boxplots_Dom

Q1.SC.length_Nested_boxplots_Musc <- ggplot(Curated_BivData99 %>% filter(subsp == "Musc") %>% filter(strain != "CZECH"), aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none") 
Q1.SC.length_Nested_boxplots_Musc

Q1.SC.length_Nested_boxplots_Mol <- ggplot(Curated_BivData99 %>% filter(strain == "MSM"), aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+geom_jitter(alpha=.3)+
  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(strain~sex, scales="free")+ylim(c(25,255))+theme(legend.position="none") 
Q1.SC.length_Nested_boxplots_Mol

The full mFull data set of SC lengths shows that

- there is a positive correlation between mean SC length and number of COs per chrm

@@

@@

Chromosome size effect.

@@

```{r Q1.CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}

@@

#write.table(round(summry.glm.Q1_IFD.ABS.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.ABS.M3.csv", sep="\t",row.names = TRUE)

@@

#write.table(round(summry.glm.Q1_IFD.PER.M3$coefficients,6), #"~./MLH1repo/summry.glm.Q1_IFD.PER.M3.csv", sep=",",row.names = TRUE)

@@

First we examine the pattern of SC length variation across chromosome classes.

```{r Q2.show.SC.length, echo=FALSE, warning=FALSE}

Curated_BivData99 <- Curated_BivData[Curated_BivData$hand.foci.count < 5,]
Curated_BivData99 <- Curated_BivData99[(!is.na(Curated_BivData99$sex)),]

#adjust order of plot here
Curated_BivData99$strain <- factor(Curated_BivData99$strain,levels =c(
  "WSB","G","LEW","PERC",  
"PWD", "SKIVE",  "KAZ",
     "MSM", "MOLF",   
                     "CAST",
                         "SPRET","SPIC" ), order=T )

Q2.SC_BivData <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)

#nested boxplots, pooled Bivalents
#x axis for hand.focicount -- good for seperating the points --- but it pools

#prediction for Dom strains: low variation 
Q2.SC.length_Nested_boxplots_Dom <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=strain, fill=as.factor(hand.foci.count) ))+geom_boxplot(alpha=0.5)+ geom_jitter(alpha = .1)+
  ggtitle("Bivalent SC Lengths by Chrm Class \n Dom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") + scale_color_manual(values=color_chrm_class)

Q2.SC.length_Nested_boxplots_Dom2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Dom",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+ geom_jitter(alpha = .1)+
  ggtitle("Bivalent SC Lengths by Chrm Class \nDom")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
Q2.SC.length_Nested_boxplots_Dom2

Q2.SC.length_Nested_boxplots_Musc2 <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Musc",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot( aes(x=strain), alpha=0.5)+geom_jitter(alpha=.2)+
  ggtitle("Bivalent SC Lengths by Chrm Class \nMusc")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none") +scale_color_manual(values=color_chrm_class)
Q2.SC.length_Nested_boxplots_Musc2

Q2.SC.length_Nested_boxplots_Mol <- ggplot(Q2.SC_BivData[Q2.SC_BivData$subsp == "Mol",], aes(y = chromosomeLength, x=as.factor(hand.foci.count), fill=as.factor(hand.foci.count) ))+geom_boxplot(aes(x=strain),alpha=0.5)+geom_jitter(alpha=.3)+  ggtitle("Bivalent SC Lengths by Chrm Class\n Mol")+facet_wrap(subsp~sex, scales="free")+ylim(c(25,160))+theme(legend.position="none")+scale_color_manual(values=color_chrm_class)
Q2.SC.length_Nested_boxplots_Mol

**general pattern: SC lengths correlate with number of foci on single bivalents (the chromosome classes). Diviations within strains are due to low sample numbers.

The positive correlation of sc lengths has a positive correlation with the number of COs on single chromosomes ,

in the musculus and molossinus strains, the positive correlation is more

In Domisticus the -- postive correlation only holds for the first 3 classes of bivalents (0CO, 1CO, 2CO) -- this illustrates how the differences between proportions of chromosome classes across strains confonds comparisons of SC length across strains.

The glm's for the full dataset across -- the coefficients are listed below

@@

We purpose -- we think that running the models for comparing SC means by chromosome class and devided by rec group. (since the chromsome porportions are similar -- hopefully/we assume the the sets of chromosomes within each set are more similar )

The null hypothesis is that these won't be highly significant strain effects.
If there are still strain effects fo these models, that would be evidence that there has been eevolution for SC length compaction across these strains

```{r show.tot.sc, echo=FALSE}

totSC.mouse.av.table <- totSC.mouse.av.table %>% filter(strain!="CAST" & strain !="F1"& strain !="SPIC" & strain !=  "CAROLI" & strain !="SPRET"& strain !="HMI"& strain !="PERC")

tot.sc99 <- ggplot(data= totSC.mouse.av.table, aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(sex~subsp, scales="free")

tot.sc77 <- ggplot(data= totSC.mouse.av.table[totSC.mouse.av.table$sex == "male",], aes(y = mean.skel, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ylim(c(150, 650))+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
tot.sc77

tot.sc_bin <- ggplot(data= totSC.mouse.av.table[totSC.mouse.av.table$sex == "male",], aes(y = mean.bin, x=mouse, color=strain))+ geom_point()+scale_color_manual(values=colors_of_strains)+
  facet_wrap(Rec.group~subsp, scales="free")+ theme(axis.title.x=element_blank(),        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())  +ylim(c(0, 3000))
```

The skeletonized measures show that the general pattern holds for total SC and the high rec strains. However there are a few exceptions -- However in the molf strain, there might be strain specific evolution--the mol subsp all have more SC skel

--

-the SKIVE are within dom and kaz level -- so why does SKIVE have inntermediate CO proportions and gwRR

24sep14_PWD_m1 is lower than the others
5jul17_MSM_m1 is the low MSM

The CZECH and AST are higher than KAZ

```{r Q2.SC_class.rec.group_setup, echo=FALSE}

#rerun the models for datasets -- deviding by chrm class and rec group
Curated_BivData$Rec.group <- ifelse(grepl("PWD male", Curated_BivData$category), 1, 
                            ifelse(grepl("MSM male", Curated_BivData$category), 1,
                             ifelse(grepl("SKIVE male", Curated_BivData$category), 1, 0)) )

Q2.SC_BivData_low.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5)%>% filter(Rec.group == 0 )
Q2.SC_BivData_high.rec <- Curated_BivData %>% filter(sex=="male") %>% filter( hand.foci.count < 5) %>% filter(Rec.group == 1 )

#calq_mouse av
mouse.av.SC.low.rec <- ddply(Q2.SC_BivData_low.rec, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                        n_cells = length(unique(fileName) ),
                        mean.SC.total = mean(chromosomeLength),

                    mean.SC_0CO = mean(chromosomeLength[which(hand.foci.count == 0)], na.rm = TRUE),
                  n_0CO =  sum((hand.foci.count == 0), na.rm = TRUE ),

                   mean.SC_1CO = mean(chromosomeLength[which(hand.foci.count == 1)], na.rm = TRUE),
                    n_1CO = sum((hand.foci.count == 1), na.rm = TRUE),

                   mean.SC_2CO = mean(chromosomeLength[which(hand.foci.count == 2)], na.rm = TRUE),
                   n_2CO = sum((hand.foci.count == 2), na.rm = TRUE),

                  mean.SC_3CO = mean(chromosomeLength[which(hand.foci.count == 3)], na.rm = TRUE),
                  n_3CO = sum((hand.foci.count == 3), na.rm = TRUE)

mouse.av.SC.low.rec <- add_strain(mouse.av.SC.low.rec)
mouse.av.SC.low.rec <- add_subsp(mouse.av.SC.low.rec)

mouse.av.SC.low.rec$strain <- factor(mouse.av.SC.low.rec$strain, ordered = FALSE)
mouse.av.SC.low.rec$subsp <- factor(mouse.av.SC.low.rec$subsp, ordered = FALSE)

#calq_mouse av
mouse.av.SC.high.rec <- ddply(Q2.SC_BivData_high.rec, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength)
                         )
mouse.av.SC.high.rec <- add_mouse(mouse.av.SC.high.rec)
mouse.av.SC.high.rec <- add_strain(mouse.av.SC.high.rec)
mouse.av.SC.high.rec <- add_subsp(mouse.av.SC.high.rec)

mouse.av.SC.high.rec$strain <- factor(mouse.av.SC.high.rec$strain, ordered = FALSE)
mouse.av.SC.high.rec$add_subsp <- factor(mouse.av.SC.high.rec$subsp, ordered = FALSE)
```

The number of cells and mice is difference across the low group strains, 
-- permut the data, sample just 3 mice per category, which have similar numbers of cells represented (or down sample)

```{r Q2.SC.Length.stuff, echo=FALSE}
#all pooled SC lengths
#WSB     G   LEW  MOLF   KAZ CZECH 
#    5     9     5     4     3     3

#1apr16_G_m4, fewest obs and fewest cells, 12oct15_G_m1 7 cells, others have ~14

#G sample 5 mice

#break down of mice in the low rec dataset

#glm.1CO
Q2.glm_SCLength_low.rec_1CO <- glm(mean.SC_1CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
summry.Q2.glm_SCLength_low.rec_1CO <- summary(Q2.glm_SCLength_low.rec_1CO)
round(summry.Q2.glm_SCLength_low.rec_1CO$coefficients[,c(1,4)],5)

#glm.2CO
Q2.glm_SCLength_low.rec_2CO <- glm(mean.SC_2CO ~ subsp * strain, data=mouse.av.SC.low.rec)
summry.Q2.glm_SCLength_low.rec_2CO <- summary(Q2.glm_SCLength_low.rec_2CO)
round(summry.Q2.glm_SCLength_low.rec_2CO$coefficients[,c(1,4)],5)
#fewer significant differences

Q2.glm_SCLength_low.rec.M2_2CO <- glm(mean.SC_2CO ~  strain, data=mouse.av.SC.low.rec)
sum.Q2.glm_SCLength_low.rec.M2_2CO <- summary(Q2.glm_SCLength_low.rec.M2_2CO)
round(sum.Q2.glm_SCLength_low.rec.M2_2CO$coefficients[,c(1,4)],5)
#all but KAZ are slightly significant, G and CZECH more so

#glm.3CO
Q2.glm_SCLength_low.rec_3CO <- glm(mean.SC_3CO ~ subsp * strain, data=mouse.av.SC.low.rec) 
#not sig differences
summry.Q2.glm_SCLength_low.rec_3CO <- summary(Q2.glm_SCLength_low.rec_3CO)
round(summry.Q2.glm_SCLength_low.rec_3CO$coefficients[,c(1,4)],5)

#Q2.lmer_SCLength_low.rec_R.sex <- lmer(mean.SC ~ subsp + (1|strain), data=mouse.av.SC.low.rec, REML = TRUE)

### 2020-02-14 doc/totalSC.Rmd
starting total SC analysis

#load BivData
load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData.RData")

@@

The mixed models are mouse averages, get mouse names and ncells for each of the lists.

@@

```{r more.blahs, echo=FALSE}

#WSB -- WSB_curation, WSB_female output on big External
#not sure why Curated bivData catching more categories than this above

@@

#output: matched random names with real names.
#save RData file
#save.image("~./MLH1repo/data/total_SC_1.6.20.RData")

>>>>

#load BivData
load(file="C:/Users/alpeterson7/Documents/MLH1repo/data/CleanBivData_13.2.20.RData")

@@

@@

```{r more.blahs, echo=FALSE}
#WSB -- WSB_curation, WSB_female output on big External
#not sure why Curated bivData catching more categories than this above

@@

#output: matched random names with real names.
#save RData file
#save.image("~./MLH1repo/data/total_SC_14.2.20.RData")

#changing to just loading the csv file
